Extending the WebID Protocol with Access Delegation

### 2012-08-02 paper.tex
added 3 step secretary relation part

\begin{abstract}
The WebID protocol enables the global identification of agents in a distributed manner by employing asymmetric cryptography and Linked Data principles.
WebIDs are mostly used in scenarios where a user utilize a browser application to access a protected resource on a web server.
In order to decide whether access should be granted or denied, the web server itself needs to retrieve other WebID profiles and resources which contain relationship information between agents and other related information.
%Currently, those information needs to be publicly available, which leads to an privacy issue, since agents may want to hide such information from the general audience.
In this paper we propose an extension to the WebID protocol which allows for delegation of access authorization from a WebID to a third party, e.g. a server application.
This extends the range of application scenarios where WebID authentication can be efficiently deployed to all kinds of server to server resource consumption scenarios and other use cases.  
\end{abstract}

@@

We can easily think of the origin server as not only able to respond to requests, but also as an agent able to make requests.
Indeed WebID authentication requires the server to make WebID profile requests to other servers in order to verify the identity of agents making a request to it.
The WebID spec describes this task as being accomplished by a separate agent, the WebID verifier - which could indeed be done by another service on the web (WebID proxy authentication servers).
But it can also be so closely tied to the web service that it would be natural to think of it as part of that same service.
The WebID profile furthermore could be served by the same agent as the one making the request, in which case we have a minimal case of a peer to peer communication.
Note that fetching a WebID profile for WebID authentication should be done anonymously, for fear of authentication deadlocks.\todo{ST: why that? done -> allowed}

Things get more interesting in the authorization space.
Consider a very natural application of WebID: allowing friends of one's friends (foaf) access to some resources.
This authorization rule will require the web server to fetch each of its users' friends profiles, in order to build up such a list.
But there is a privacy issue involved here: not every one wants to make all of their social network (or even any of it) publicly visible.
Those people may then protect their FOAF profile with access control rules.
How can a server that needs access to these FOAF profiles in order to apply its own access control rules get access to the information?
Should it be able to do so?
These are the questions we will try to answer in this paper.

@@

\textit{Solution 1: Using each users WebID} vs.
\textit{Solution 2: Acting on Behalf of a User}

@@

\subsection{MyProfile description}
MyProfile intends to provide a solution for managing the numerous accounts and profiles that users have on the Internet. Its main purpose is to provide a unified user account, or simply \textit{user profile}, which as opposed to current \textit{silo} profiles, would really be under the user's control, on a device located within the user physical reach.

In the case of MyProfile, it is very important to be able to offer WebID access delegation, because a single MyProfile server instance can host multiple users. To improve user experience and overall performance, a caching mechanism is used to refresh local copies or "views" of external data. Due to multiple users coexisting on the same server, the caching mechanism needs to be able to differentiate between views belonging to each user, according to their specific access control policies configured on external resource providers.

Let's take for example the following case: Ann and Barry are both local users on a single MyProfile server. Charley is an external user with access control policies set up on his private server. When Ann requests to view Charley's profile data, a personalized view of the profile is displayed, corresponding to access control policies for Ann. When Barry requests to view Charley's profile, different profile data is displayed, since there are different access control policies for Barry. At this point, the caching mechanism needs to be able to cache two different profile views, each corresponding to access control policies specific for the user requesting the data. For this reason, the caching mechanism is required to act as a service agent, identifying itself as well as conveying the identity of the user on  behalf of whom the agent is requesting data.

@@

%A fix for cases where there is little trust in a secretary, or when bootstrapping a user's server. The resource provider can store the request from consumer, and send a pingback to the consumer service, to notify him/her of a pending authorization request from consumer. Then the user will have to authenticate personally (through the browser, in order to verify the private key) to the provider, check a list of pending requests for his/her WebID, and decide if the request in question is valid or not. In other words, the user must give his/her consent if the secretary's identity is unknown to the provider.

>>>>

\begin{abstract}
The WebID protocol enables the global identification of agents in a distributed manner by combining asymmetric cryptography and Linked Data.
WebIDs are mostly used in scenarios where a user utilize a browser application to access a protected resource on a web server.
But in order to decide whether access should be granted or denied, the web server itself needs to retrieve other WebID profiles and linked resources in order to work out if the requesting agent is part of the authorization group ( eg: is a friends of a friend of the resource owner ).
It would seem that at present this information needs to be publicly available for the server to access it. This  leads to a privacy issue, since agents may want to limit access to a users social graph.
In this paper we propose an extension to the WebID protocol which allows for delegation of access authorization from a WebID to a third party, e.g. allowing a server to be able to act on behalf of its users.
This extends the range of application scenarios where WebID authentication can be efficiently deployed while increasing privacy.
\end{abstract}

@@

We can easily think of the origin server as not only able to respond to requests, but also as an agent able to make requests.
Indeed WebID authentication requires the server to make WebID profile requests to other servers in order to verify the identity of agents making a request to it.
The WebID spec describes this task as being accomplished by a separate agent, the WebID verifier - which could indeed be done by another service on the web (WebID proxy authentication servers).
But it can also be so closely tied to the web service that it would be natural to think of it as part of that same service.
The WebID profile furthermore could be served by the same agent as the one making the request, in which case we have a minimal case of a peer to peer communication.
Note that fetching a WebID profile for WebID authentication should be done anonymously, for fear of authentication deadlocks.\footnote{ for example one can imagine an agent S  with profile Ps requesting a resource on server R which requires authentication. S would send R its certificate, thereby requiring  R to dereference S's Profile Ps in order to verify the WebID. If Ps  itself requires authentiction of R and if R sends a certificate containing a WebID with its profile Pr, and if Pr itself requires authentication then it looks like we have a deadlock.  }

Things get more interesting in the authorization space.
Consider a very natural application of WebID: allowing friends of one's friends (foaf) access to some resources.
This authorization rule will require the web server to fetch each of its users' friends profiles, in order to build up such a list.
But there is a privacy issue involved here: not every one wants to make all of their social network publically visible, and some may not want to make any of it publicly visible.
Those people may then protect their FOAF profile with access control rules.
How can a server that needs access to these FOAF profiles in order to apply its own access control rules - such as only allowing access to foaf - get access to the information?
Should it be able to do so?
These are the questions we will try to answer in this paper.

@@

\textit{Solution 1: Passing itself for the user}

   The simplest solution for any agent S wishing to pass itself off as one of the users U, is for it to create a public/private key pair,  place the public key in a certificate Cu and add the public key to the Users' Profile Pu. Having done this the agent S can then connect using the certificate Cu to any service it wishes to, when it wants to work for the user U. This has the advantage of requiring no change to the WebID protocol. It does give the agent S all powers of the user U and so has to be handled with great care.

  What kind of agents does it make sense to give such rights to? Or put another way: what agent would giving such rights to not constitute a reduction in security to the user? Or put yet another way: which agent could take such a role at its behest at any moment in any case? Or: which agent does the user and the authorization server already rely on for authentication? There is one agent  on which any service requiring WebID authentication relies on and this is the WebID Profile Origin Server \footnote{see http://tools.ietf.org/html/rfc6454}. Any server requiring authentication of an agent using WebID must trust the origin server serving the WebID Profile. That origin server can at any time change the public keys in the profiles sent back to any authorization agents and indeed change any information in the profile it wants to. Any user profile can then be thought of as being a user according to the Origin Server ie a user@origin.org:port - where origin.org is the domain of the origin server, and 'port' is the port the server is running on.

  When the user owns the origin server then the origin server is just an extension of the user, and in that case there need be no thought of a difference between the user and the origin server. In that case the origin server can in fact just authenticate as the user whenever needed.

  But when the origin server (OS) is serving a number of different profiles and is acting on behalf of a number of different users, then we have an efficiency issue that rears its head. Whenever the Origin Server needs to act on behalf of one ifs users to a remote server R it will need to open a new TLS connection to that server using one of its users public keys. For N users this could require N parallel connections to a server S.

\todo{calculate how inefficient this is, and at what size of network this starts really become problematic}

\textit{Solution 2: Origin Server acting on Behalf of a User}

  In order to reduce the number of open connections to any server down to a minimum, it would be useful if the Origin Server could identify itself directly when making a request and specify on behalf of which of its users it was acting on. 
One way would be for the Origin Server acting as client to simply use the same public key as the Origin Server acting as server - that is it could use the same public key as the one used by the TLS server.  
It could even use the same certificate. 
This would make clear to any server it was connecting to, that it could rely on this server as much as it could on the identity of any user served by that server.

Still, such a server connecting to a remote service on behalf of one of its users would then need to identify which user it was acting on behalf of, or the remote service receiving a request would not know which access control rule to apply to the origin server.

After all the remote service may have resources which should be readable by the Origin Server users OSu1 and OSu2, but not to the Origin Server users OSu3 and OSu4. Or the representations returned may differ somewhat for each group of users. OSu1 and OSu2 may get some more information than OSu1 and OSu2. As a result it is important for the Origin Server to distinguish on Behalf of which users it is acting when making a request.

Problems: using the same key between server acting as client and as server risks making the keys available to a wider audience, and so increases the risk of key compromise.

\todo{insert description of On-behalf-Of header}

\textit{Solution3: Secretary acting on Behalf of a User}

The origin server acting  as a client on behalf of a user can then be thought of as a keep of secrets for that user. 
It should know how to distinguish what remote servers tell it when it is acting on behalf of one user, from what a remote server tells it when it is acting on behalf of another user. The role of the keeper of secrets for a person is known as the secretary role - a prestigious role taken on for example by figures such as the Secretary of State, Hillary Clinton.

If we now identify the secretary that can act on behalf of a user we can generalise the protocol somewhat. When the server authenticates as a secretary to a remote server it can use its own WebID. But how would the remote server know that it can trust that secretary to be acting on behalf of a particular user? It can no longer just compare the TLS keys. Insterad it can see if the user that the secretary is acting on behalf of has stated clearly that that agent is its secretary. This can be discovered by checking out the profile... \todo{fill out the details here}

@@

\todo{Andrei Sambra: MyProfile description}

@@
